### 2012-03-05 papers/eccv2012/main.tex
ECCV review submission

\begin{abstract}
  We consdier the problem of learning to reconstruct polygonal models
  from single and multiple views.

  In this difficult domain we start small and focus on indoor
  Manhattan models.

  We learn lots about practical structured prediction in this unique
  domain

  We kick the state-of-the-art's butt on everything :)
\end{abstract}

@@

Reconstruction long considered a central problem in computer vision

Multiple view reconstruction long history, mostly uses probabilistic
models with few free parameters

Single view reconstruction a recent but now well-established
trend. Learning obviously a core part of this fundamentally
under-determined problem. But not much care on loss functions,
optimality guarantees, or statistical consistency.
-> Exceptions: Hedau et al (but very simple models)

Structured prediction used for increasingly complex models. Its use in
computer vision part of a long trend towards statistically rigorous
and well-understood convex optimization techniques for solving
empirical risk minimization problems in computer vision.

Has been used much before for reconstruction. Stereo for example, and
Hedau. Here we go all the way to an expressive high-level geometric
model class.

Attraction of doing this is many fold:
Can minimize model-level loss rather than pixel-level loss
-> give a nice demonstration of this
Can minimize realistic loss functions in a rigorous way
Can learn in both single and multiple view setting using the *same*
model
Bayesian inference algorithm

We learn much about the practice of structured prediction of value to
the community and focus an entire section of conveying these lessons.

Out approach: describe the hypothesis class, describe the model

The remainder of this paper is...

@@

We're drawing together multiple view reconstruction, single view
reconstruction, and structured prediction -> cannot give a reasonable
survey, sorry!!

Reconstruction papers

Multi view reconstruction, stereo -> not much learning

Single view -> learning but mostly at pixel level. Some exceptions

Learning papers

Structured prediction, reconstruction applications

@@

We now describe the indoor Manhattan model that we are attempting to
learn. As with any learning problem, there are three ingredients that
we need to specify: a hypothesis class, an observation model for
hypotheses, and a loss function. In our setup these are, respectively,
the class of indoor Manhattan models, a log--linear Bayesian
likelihood, and either the relative depth error or a labelling error
(we describe both).

@@

This paper is concerned with the hypothesis class consisting of indoor
Manhattan reconstructions, which are 3D polygonal models characterized
by infinite floor and ceiling planes with vertical walls extending
between them \cite{Lee09,Flint10cvpr}. Indoor Manhattan environments
are a sub--class of general Manhattan environments. Examples are shown
in figure XXX.

This is an attractive hypothesis class because
\begin{enumerate}
\item{it captures many \textit{regularities} within man--made
  environments;}
\item{the geometric primitives (floor/wall/ceiling) are immediately
  useful for semantic--level scene understanding;}
\item{it is expressive enough to represent approximately or exactly
  a surprisingly wide variety of environments;}
\item{there is a simple and convenient parametrisation;}
\item{an efficient inference algorithm exists \cite{Flint10eccv}.}
\end{enumerate}

An indoor Manhattan model is much like an archicetural floor--plan,
and can be specificied as a set of 2D line segments representing walls
together with the position of the floor and ceiling plane (figure
XXX). In this paper we adopt an image--domain parametrisation due to
its convenience for inference and learning. Following Flint \etal
\cite{Flint11} we represent hypotheses as a path running from the left
edge to the right edge of the image, which we will refer to as a
\textit{seam}. A seam $\Seam$ consists of a sequence of pairs of
scalars,
\begin{equation}
  \Seam = \{\seam_i,\Label_i\}_{i=1}^{\Width}
\end{equation}
where $\seam_i$ is the $y$--cordinate at which the path intersects
image column $i$, $\Orient_i$ is the orientation of the wall in that
column, and $\Width$ is the image width. Remarkably, this simple
parametrisation specifies a unique metric 3D model up to scale
\cite{Flint11}.

While we do not here have space for a full discussion of the geometry
of indoor Manhattan models, there are two properties of the seam
representation that will be of relevance to the following
sections. Firstly, images of indoor Manhattan environments can be
rectified so that vertical lines in the world project to vertical
lines in the image \cite{Flint10eccv}. Secondly, indoor Manhattan
models decompose in the following manner. We said earlier that a seam
$\Seam$ specifies a unique 3D reconstruction. Let
$\Orient(x,y;\Seam)\in\TheOrientations$ be the orientation of the 3D
surface projecting to pixel $(x,y)$. Then Flint \etal
\cite{Flint10eccv} showed that $\Orient$ is functionally dependant
\textit{only} on the labels for column $x$. That is, we may write
\begin{equation}
  \label{eq:orient-indep}
  \Orient(x,y;\Seam) = \tilde{\Orient}(x,y;\seam_x,\Label_x) ~.
\end{equation}
Similarly, letting $\Depth$ be the depth of the reconstruction at
$(x,y)$ we may write $\Depth$ in terms of a single label pair,
\cite{Flint10iccv}
\begin{equation}
  \label{eq:depth-indep}
  \Depth(x,y;\Seam) = \tilde{\Depth}(x,y;\seam_x,\Label_x) ~.
\end{equation}

\subsection{Observation Model}

We adopt the probabilistic model formulated by Flint \etal
\cite{Flint11}, which relates indoor Manhattan hypotheses to
single--view image features, multiple--view photoconsistency terms,
and a reconstructed point cloud. From our perspective these are simply
features and any subset may be omitted if in appropriate for a given
application.

TODO: walk through the graphical model?

Under this model the prior on environments is a geometric distribution
in the number of distinct wall segments, which means, roughly, that we
consider reconstructions with more walls apriori less likely. The
benefit of this prior is that its logarithm is linear in the number of
wall segments,
\begin{equation}
  \log P(\Seam) = - \vect{n} \cdot \Penalties
\end{equation}
where $\vect{n}$ is a vector containing the number of walls in $\Seam$
of various categories (concave, convex, occluding) and $\Penalties$ is
a vector of hyper--parameters.

Flint \etal showed that for a variety of single--view and
multiple--view features $\Features$ there are reasonable choices of
likelihood that can be written in the form
\begin{equation}
  \log P(\Features ~|~ \Seam) = \sum_{x=1}^\Width 
  \PixelPayoff_{\Features}(x,\seam_x)
\end{equation}
where $\PixelPayoff_{\Features}$ is a 2--dimensional matrix computed
deterministically from $\Features$ \cite{Flint11}. It is easy to show
that for each likelihood in \cite{Flint11}, the log--likelihoods
is linear in the hyper--parameters $\LikelihoodHypers$,
\begin{equation}
  \log P(\Features ~|~ \Seam) = 
  \sum_{x=1}^\Width
    \LikelihoodHypers_{\Features} 
    \cdot
    \PixelFtr_{\Features}(x,\seam_x)
\end{equation}

Assuming conditional independence between the feature types, we have
for the posterior,
\begin{eqnarray}
  P(\Seam ~|~ \Features_1 \ldots \Features_n)
  &\propto&
  P(\Seam) P(\Feature_1 ~|~ \Seam) \ldots P(\Feature_n ~|~ \Seam)\\
  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
  &=&
  - \vect{n} \cdot \Penalties
  + \sum_{i=1}^n \sum_{x=1}^\Width 
      \LikelihoodHypers_i
      \cdot
      \PixelFtr_i(x,\seam_x)
  + O(1) ~.
  \label{eq:log-posterior}
\end{eqnarray}

In order to place our model within a structured prediction framework
we need to re--write the above in terms of a joint feature space
$\JointFtr(\Seam,\Features)$. The decomposability of indoor Manhattan
mdoels into payoff matrices permits precisely such a
formulation. Defining
\begin{align}
  \JointFtr(\Seam,\Features) =
  \begin{bmatrix}
    -\vect{n} \\
    \sum_{x=1}^\Width \PixelFtr_1(x,\seam_x) \\
    \vdots \\
    \sum_{x=1}^\Width \PixelFtr_n(x,\seam_x)
  \end{bmatrix}
  & \quad &
  \Model =
  \begin{bmatrix}
    \Penalties \\
    \LikelihoodHypers_1 \\
    \vdots \\
    \LikelihoodHypers_n
  \end{bmatrix}
  \label{eq:joint-ftr-def}
\end{align}
we see that \eqnref{log-posterior} can be written
\begin{equation}
  \log P(\Seam ~|~ \Features_1 \ldots \Features_n)
  =
  \bigl\langle \Model ~,~ \JointFtr(\Seam,\Features) \bigr\rangle + O(1)
\end{equation}
where we have adopted Hilbert space notation with
$\langle\cdot,\cdot\rangle$ denoting an inner product. Since $\Model$
contains all the free parameters in the model, the goal learning will
be to optimise $\Model$ with respect to a set of training features
$\{\Features_i\}$ together with ground truth labellings $\{\Seam_i\}$.

@@

The precise make--up of the feature space depends on the particular
forms for the likelihoods and which sensor modalities are
available. We define separate feature spaces for the single-- and
multiple--view contexts; both are summarized in \figref{featurespace}.

\begin{figure}[tb]
  \centering
  \begin{tabular}{@{}llll@{}}
    \toprule
    Feature & Dimensionality & Single--view? & Reference \\
    \midrule
    Line sweeps & 1 & yes & Lee \etal \cite{Lee09}\\
    Stereo photo--consistency & 4 & no & Flint \etal \cite{Flint11}\\
    Point cloud & 2 & no & Flint \etal \cite{Flint11}\\
    \bottomrule
  \end{tabular}
  \caption{The composition of our feature space. Each row corresponds
    to one or more of the $\PixelFtr_i$ in \eqnref{joint-ftr-def}.}
  \label{fig:featurespace}
\end{figure}

@@

A loss function $\Delta(\Seam,\EstSeam)$ is a measure of the
deviation of a predicted label $\EstSeam$ from the ground truth
$\Seam$. In the context of learning one often faces a trade--off
between choosing a loss that leads to tractible optimization, and
choosing a loss that measures what one ``really'' cares about. For
example, Hoiem \etal \cite{Hoiem05} implicitly use a per--pixel loss,
since this decomposes optimization pixel--wise. However, it would be
hard to argue that pixel--wise accuracy is an honest measure of what
the authors care about: some pixel--level errors are insignificant to
the overall quality of a reconstruction, while others are
catastrophic. Whether the authors care about the accuracy of their
reconstruction in 3D, its utility for contextual object detection, or
even the aesthetic appeal of reconstructions, a per--pixel loss is
almost certainly a weak measure of success.

A more reasonable choice would be the relative depth error, which has
been the gold standard within the reconstruction community for decades
\cite{Hartley04}, and measures the average deviation between
reconstructed and ground truth depths:
\begin{equation}
  \DepthLoss(\Seam,\EstSeam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
    \frac
        {\Depth(\Pixel;\Seam) - \Depth(\Pixel;\EstSeam)}
        {\Depth(\Pixel;\Seam)} ~,
\end{equation}
where $N$ is the number of pixels. Another reasonable choice is the
labelling error, used widely within the semantic segmentation
literature,
\begin{equation}
  \LblLoss(\Seam,\EstSeam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
    \bigl[~\Orient(\Pixel;\Seam) \neq \Orient(\Pixel;\EstSeam)~\bigr] ~,
\end{equation}
where $[p]$ is 1 if $p$ is true and 0 otherwise. An attractive
characteristic of the indoor Manhattan class is that both of these
loss functions can be optimized exactly. The algorithmic details for
this are left to \sectref{learning}; the key result that we need to
attend to here is that $\DepthLoss$ and $\LblLoss$ can be written in a
form similar to the payoff formulation for the feature
likelihoods. Recall the functional independence established in
\eqnref{func-indep}.

In the
case of reconstruction, the gold standard is the relative depth error
\cite{Hartley04}

Should we show how to write these as payoff matrices here, or leave
this to the learning section?

L1, L2 error: easy representation, not meaningful, give examples

Labelling error
-> showing these will require some of the ICCV framework to have been presented
-> give nice intuition, show scatter plots, examples

Relative depth error
-> showing these will require some of the ICCV framework to have been presented
-> give nice intuition, show scatter plots, examples

None of these are pixel classifiers!

Writing loss functions in the seam representation -> maybe just
mention that this will be useful later

\section{Inference}
\label{sec:inference}

Describe the inference algorithm in terms of payoff/penalty
optimization, cite ECCV -> this section must be short!

@@

Risk minimization formulation
%  -> min_f E[ Loss(f(x),y*) ]
Empirical risk problem
%  -> min_f sum_{xi} Loss(f(x),y*)
Write f as a linear argmax
%  -> f(x) = argmax_y <w,psi(x,y)>
Reduced to quadratic optimization (cite SVMstruct papers)
%  -> min_f ...
-> subject to ....
Minimize using algorithm in ...
This requires a separation oracle, which is the topic of the
following subsection

Loss-augmented inference
-> the problem
-> writing loss the functions as payoff matrices
-> writing a payoff matrix for loss-augmented inference
-> write a proof that ``the optimal model payoff matrix
B=... solves the loss-augmented inference problem <w,psi> + L

\section{Practical Considerations}
\label{sec:practical-considerations}

- negative slacks are more harmful than one would expect

- inconsistent inference is very harmful

- inference when the ground truth not in the hypothesis class

- conditioning in a joint feature space

>>>>

\begin{abstract}
  We develop a structured prediction approach to reconstructing 3D
  polygonal models from single and multiple images of a
  scene. Building on recent advances in single view reconstruction we
  adopt the indoor Manhattan hypothesis class --- one of the most
  complicated output spaces (in terms of internal constraints) yet
  considered within a structured prediction framework. Our approach
  can learn in both the single-- and multiple--view contexts. We
  show that the chosen hypothesis class permits optimizing a variety
  of high--level loss functions, such as the relative depth error. Our
  results out--perform the state--of--the--art, including an
  improvement of more than $50\%$ on one metric.
\end{abstract}

@@

Reconstruction 3D models from images is a central problem in computer
vision. There is a deep literature concerning reconstruction from
multiple views of a scene, with particular focus on the geometry of
multiple cameras. More recently the recovery of 3D information from a
single image has received attention \cite{Hoiem05,Saxena09}. In this
context the geometric constraints leveraged by multiple--view
reconstructors are unavailable, so inference must rely on photometric
cues. To capture the complex, high--dimensional patterns involved in
this challenging inference task, practitioners have leveraged a range
of machine learning techniques.

Within single--view reconstruction, few authors have cast learning as
a single optimisation with respect to a clearly defined loss function
\cite{Hoiem05,Saxena09,Lee09,Flint11}, while most approaches to
multiple--view reconstruction do not consider learning from training
data at all. In contrast, this paper casts reconstruction
fundamentally as a learning problem, with the goal being to learn a
prediction function $\Predictor$ mapping observed features to 3D
reconstructions.

Building on recent advances in single--view reconstruction we adopt as
our hypothesis class the set of indoor Manhattan models
\cite{Lee09,Flint10eccv,Flint11}, under which scenes are approximated
by a floor and ceiling plane together with a sequence of vertical
walls. This representation brings a variety of attractive features
such as a simple parametrisation, efficient and exact inference,
decomposability of loss functions, and a balance between
expressiveness and robustness.

For learning we employ the tools of structured prediction, in
particular the structural SVM \cite{Tsochantaridis04}. The use of
these tools is part part of a long trend towards statistically
rigorous, well--understood convex optimization techniques in computer
vision. Recent successfully applications include detection
\cite{blaschko2008learning}, segmentation \cite{taskar2004max}, and
scene classification \cite{Yang2010}. In the domain of reconstruction,
structured prediction ideas have been applied to several simple models
classes such as stereo disparities \cite{li2008learning} and cuboids
\cite{Hedau09}.

The application of structured prediction to the indoor Manhattan class
of models constitutes one of the most complex output spaces yet
considered within this framework. The indoor Manhattan model enforces
hard geometric constraints that lack simple expressions in terms of
image coordinates. These constraints are context--dependent, being
tied to quantities such as camera rotation and the location of
vanishing points. We have learnt several valuable lessons of general
relevance from this complex prediction task, to which we dedicate the
final section of this paper.

The contributions of this paper are thus (i) a unified learning
framework for single-- and multiple--view reconstruction, utilising
the indoor Manhattan model and the tools of structured prediction;
(ii) the reduction of two image--level loss functions to a form
amenable to efficient optimization; (iii) an efficient separation
procedure for identify the ``most--violated constraint'' during
learning, (iv) an empirical demonstration of structured prediction in
perhaps the most complex output space yet considered within this
framework; and (v) a series of practical observations concerning the
application of structured prediction techniques.

In the remainder of this paper we present background material
(\secref{background}), followed by the indoor Manhattan model itself
(\secref{model}), and our learning framework (\secref{learning}). We
then present results for multiple--view reconstruction
(\secref{mv-results}) followed by single--view reconstruction
(\secref{sv-results}), then we close with practical lessons learnt
(\secref{discussion}) and concluding remarks (\secref{conclusion}).

@@

This paper touches on several major research areas (multiple--view
reconstruction, single--view reconstruction, and structured
prediction), so here we present only key contributions and those
results of particular relevance to our own work.

Multiple view reconstruction has a long history in the literature,
beginning at least as early as the seminal work of Marr \etal
\cite{Marr1976}. Low--level approaches estimate the depth of each
pixel by solving for stereo disparity (for a survey see
\cite{Scharstein2002}), while high--level approaches attempt to
recover polyhedral models of various kinds
\cite{Baillard1999,Liebowitz1999}. The present work follows the spirit
of the latter approach.

Coughlan and Yuille \cite{Coughlan99} first introduced the Manhattan
world assumption, in which reconstructed surfaces are restricted to three
mutually--orthogonal orientations. Furukawa \etal \cite{Furukawa09}
applied this idea within a multiple view stereo context. Lee \etal
\cite{Lee09} proposed the indoor Manhattan assumption, which places
further restrictions on reconstructed models.

Single--view reconstruction from line drawings also has a long history
in the literature (\cite{Huffman71}, for example). Reconstructing
single real--world images was re--introduced to the community by Hoiem
\etal \cite{Hoiem05}, who employed decision trees to classify image
segments into orientation classes in a multiple segmentation
approach. Saxena \etal \cite{Saxena09} employed energy minimisation to
recover pixel--wise depths. Barinova \etal \cite{Barinova08}
reconstructed piece--wise planar outdoor scenes using an EM
algorithm. Hedau \etal \cite{Hedau09} modelled indoor scenes as the
interior of a cuboid, and cast learning within a structured prediction
framework. We also use structured prediction, though our hypothesis
class is far more expressive.

Lee \etal \cite{Lee09} were the first to reconstruct indoor Manhattan
scenes. They devised a combinatorial search over line segments using a
branch--and--bound algorithm. Flint \etal
\cite{Flint10eccv} refined this to an exact dynamic
programming solution, and in later work \cite{Flint11} extended the
indoor Manhattan assumption to the multiple view domain. The authors
of \cite{Flint11} did describe a learning algorithm based on
bootstrapping, though its statistical consistency was given little
attention. The present work relies extensively on this thread of
research, though in contrast to past work our focus is entirely on
learning in a statistically rigorous framework. We provide comparisons
with these approaches.

Extending binary and multi--label classification to general output
spaces is a major research programme within the machine learning
community; an excellent introduction is given by Bakir \etal
\cite{Bakir07}. Our approach employs the structured support vector
machine, first proposed by Tsochantaridis \etal
\cite{Tsochantaridis04} and improved upon by Joachims \etal
\cite{Joachims2009}.

@@

In this section we describe three components of the model that we are
trying to learn (and, at test time, infer): a hypothesis class, a
feature space, and a loss function. In our setup these are,
respectively, the class of indoor Manhattan models, a log--linear
Bayesian likelihood, and either the relative depth error or a
labelling error (we describe both).

@@

This paper is concerned with the hypothesis class consisting of indoor
Manhattan reconstructions, which are 3D polygonal models characterized
by infinite floor and ceiling planes with vertical walls extending
between them (as originally proposed by Lee \etal
\cite{Lee09}). Indoor Manhattan environments are a sub--class of
general Manhattan environments; examples are shown in figure
\figref{indoor-manhattan}. This is an attractive hypothesis class
because
\begin{enumerate}
\item{it captures many regularities within man--made environments;}
\item{the geometric primitives (floor/wall/ceiling) are immediately
  useful for semantic--level scene understanding;}
\item{it is expressive enough to represent approximately or exactly
  a surprisingly wide variety of environments;}
\item{there is a simple and convenient parametrisation;}
\item{an efficient inference algorithm exists \cite{Flint10eccv}.}
\end{enumerate}

\begin{figure}[tb]
  \centering
  \subfloat[]{
    \includegraphics[width=0.3\textwidth]{figures/example-model1}
  }
  \subfloat[]{
    \includegraphics[width=0.3\textwidth]{figures/example-model2}
  }
  \subfloat[]{
    \includegraphics[width=0.3\textwidth]{figures/example-floorplan}
  }
  \caption{(a-b) Examples of indoor Manhattan environments. The red
    line illustrates the seam representation. (c)
    An indoor Manhattan environment viewed from above.}
  \label{fig:indoor-manhattan}
\end{figure}

An indoor Manhattan model is much like an archicetural floor--plan,
and can be specified as a set of 2D line segments representing walls
together with the position of the floor and ceiling plane
(\figref{indoor-manhattan}). In this paper we adopt an image--domain
parametrisation due to its convenience for inference and
learning. Following Flint \etal \cite{Flint11} we represent hypotheses
as a path running from the left edge to the right edge of the image,
which we will refer to as a \textit{seam}. A seam $\Seam$ consists of
a sequence of pairs of scalars,
\begin{equation}
  \Seam = \{\seam_i,\Label_i\}_{i=1}^{\Width}
\end{equation}
where $\seam_i$ is the $y$--coordinate at which the path intersects
image column $i$, $\Label_i$ is the orientation of the wall in that
column, and $\Width$ is the image width. Remarkably, this simple
parametrisation specifies a unique metric 3D model up to scale
\cite{Flint11}.

While we do not here have space for a full discussion of the geometry
of indoor Manhattan models, there are two properties of the seam
representation that will be relevant in the following
sections. Firstly, images of indoor Manhattan environments can be
rectified so that vertical lines in the world project to vertical
lines in the image \cite{Flint10eccv}. Secondly, the mapping from seam
to 3D reconstruction decomposes in the following manner. We said
earlier that a seam $\Seam$ specifies a unique 3D reconstruction. Let
$\Orient(x,y;\Seam)$ be the orientation of the 3D
surface projecting to pixel $(x,y)$. Flint \etal \cite{Flint10eccv}
showed that $\Orient$ is functionally dependant \textit{only} on the
pair $(\seam_x,\Label_x)$ for column $x$. That is, we may write
\begin{equation}
  \label{eq:orient-indep}
  \Orient(x,y;\Seam) = \tilde{\Orient}(x,y;\seam_x,\Label_x) ~.
\end{equation}
Similarly, letting $\Depth(x,y,;\Seam)$ be the distance from the
camera to the surface projecting to $(x,y)$ we may write
\begin{equation}
  \label{eq:depth-indep}
  \Depth(x,y;\Seam) = \tilde{\Depth}(x,y;\seam_x,\Label_x) ~.
\end{equation}

\subsection{Feature Space}

We adopt the probabilistic model formulated by Flint \etal
\cite{Flint11}, which relates indoor Manhattan reconstructions to
single--view image features, multiple--view photo--consistency terms,
and a reconstructed point cloud. From our perspective these are simply
features and any subset may be omitted if inappropriate for a given
application. Under this model the prior on reconstructions is
\begin{equation}
  \log P(\Seam) = - \vect{n} \cdot \Penalties + O(1)
  \label{eq:prior}
\end{equation}
where $\vect{n}$ is a vector containing the number of walls in $\Seam$
of various categories (concave, convex, occluding) and $\Penalties$ is
a hyper--parameter. We may interpret \eqnref{prior} as,
roughly, that reconstructions with more walls are less likely.

Flint \etal \cite{Flint11} showed that for a variety of features
$\Feature$ (including single--view and multiple--view) there are
reasonable choices of likelihood that can be written
\begin{equation}
  \log P(\Feature ~|~ \Seam) = \sum_{x=1}^\Width 
  \PixelPayoff_{\Feature}(x,\seam_x)
  \label{eq:likelihood}
\end{equation}
where $\PixelPayoff_{\Feature}$ is a real matrix computed
deterministically from $\Feature$ and $\PixelPayoff_{\Feature}(i,j)$
is the element at row $i$, column $j$. It is easy to show that for each
sensor model described in \cite{Flint11} the log--likelihood is linear
in the hyper--parameters $\LikelihoodHypers$, \ie
\begin{equation}
  \log P(\Feature ~|~ \Seam) = 
  \sum_{x=1}^\Width
  \LikelihoodHypers_{\Feature} 
  \cdot
  \PixelFtr_{\Feature}(x,\seam_x) ~.
\end{equation}
Assuming conditional independence between features, the posterior is
\begin{eqnarray}
    P(\Seam ~|~ \Features)
  &\propto&
    P(\Seam) \prod_{i=1}^n P(\Feature_i ~|~ \Seam)\\
    \log P(\Seam ~|~ \Features)
  &=&
    - \vect{n} \cdot \Penalties
    + \sum_{i=1}^n \sum_{x=1}^\Width 
    \LikelihoodHypers_i \cdot \PixelFtr_i(x,\seam_x)
    + O(1) ~.
  \label{eq:log-posterior}
\end{eqnarray}

So far, this model closely follows that described in \cite{Flint11};
our contribution is to place this model into a statistically rigorous
learning framework. To do this we need to rewrite the above in terms
of a joint feature function $\JointFtr(\Features,\Seam)$ and a
parameter vector $\Model$. The decomposability of indoor Manhattan
models into payoff matrices permits precisely such a
formulation. Defining
\begin{align}
  \JointFtr(\Features,\Seam) =
  \begin{bmatrix}
    -\vect{n} \\
    \sum_{x=1}^\Width \PixelFtr_1(x,\seam_x) \\
    \vdots \\
    \sum_{x=1}^\Width \PixelFtr_n(x,\seam_x)
  \end{bmatrix}
  & \quad &
  \Model =
  \begin{bmatrix}
    \Penalties \\
    \LikelihoodHypers_1 \\
    \vdots \\
    \LikelihoodHypers_n
  \end{bmatrix}
  \label{eq:joint-ftr-def}
\end{align}
we see that \eqnref{log-posterior} can be written as
\begin{equation}
  \log P(\Seam ~|~ \Features)
  =
  \bigl\langle \Model ~,~ \JointFtr(\Features,\Seam) \bigr\rangle +
  O(1)
  \label{eq:log-linear-posterior}
\end{equation}
where we have adopted Hilbert space notation with
$\langle\cdot,\cdot\rangle$ denoting an inner product. Since $\Model$
contains all free parameters in the model, the goal of learning will
be to optimise $\Model$ with respect to a training set
$\{(\Features_i,\Seam_i)\}$.

@@

The precise make--up of the feature space depends on the available
sensor modalities. We define separate feature spaces for the single--
and multiple--view contexts; these are summarized in
\figref{featurespace}.

\begin{figure}[tb]
  \centering
  \begin{tabular}{@{}lllll@{}}
    \toprule
    Feature & Dimensionality & Multi--view? & Single--view? & Reference \\
    \midrule
    Stereo photo--consistency & 4 & yes & no & Flint \etal \cite{Flint11}\\
    Point cloud & 2 & yes & no & Flint \etal \cite{Flint11}\\
    Line sweeps & 1 & yes & yes & Lee \etal \cite{Lee09}\\
    RGB+HSV & 6 & no & yes & \\
    Gabor responses\footnotemark & 12 & no & yes & \\
    \bottomrule
  \end{tabular}
  \caption{The composition of our single-- and multiple--view feature
    space. We omit color and Gabor features from the multiple--view
    feature space for training efficiency.}
  \label{fig:featurespace}
\end{figure}

@@

\footnotetext{4 orientations, 3 scales}

Next we define a loss function $\Delta(\Seam,\EstSeam)$ measuring the
cost of predicting some reconstruction $\EstSeam$ when in fact the
true reconstruction is $\Seam$. In the context of learning one often
faces a trade--off between choosing a loss that leads to tractable
optimization, and choosing a loss that measures the quantity that one
``really'' cares about. For example, Hoiem \etal \cite{Hoiem05} learn
a per--segment orientation classifier, then pass this as input to a
separate 3D reconstruction system \cite{Hoiem2005}. However, what one
``really'' cares about is some loss defined on the output of the
entire system rather than the output of individual components, since
some segment--level mistakes are insignificant to the overall
reconstruction quality, while others are catastrophic. This is not a
criticism of the authors' choice, but an illustration of the
trade--off faced when choosing a loss. In this paper we show how to
learn efficiently with respect to a loss defined on the final
reconstruction.

The relative depth error has been the gold standard within the
reconstruction community for more than a decade \cite{Hartley04}, and
measures the average deviation between reconstructed and ground truth
depths. In our notation,
\begin{equation}
  \label{eq:depth-loss}
  \DepthLoss(\Seam,\EstSeam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
  \frac
      {\abs{\Depth(\Pixel;\EstSeam) - \Depth(\Pixel;\Seam)}}
      {\Depth(\Pixel;\Seam)} ~,
\end{equation}
where $N$ is the number of pixels. Another reasonable choice is the
labelling error, used widely within the semantic segmentation
literature,
\begin{equation}
  \label{eq:label-loss}
  \LblLoss(\EstSeam,\Seam)
  =
  \frac{1}{N}
  \sum_{\Pixel}
  \bigl[~\Orient(\Pixel;\EstSeam) \neq \Orient(\Pixel;\Seam)~\bigr] ~,
\end{equation}
where $[p]$ is 1 if $p$ is true and 0 otherwise. An attractive
characteristic of the indoor Manhattan class is that \textit{both of
  these losses can be optimized exactly}. The algorithmic details are
left to \sectref{learning}; the key result we establish here is that
$\DepthLoss$ and $\LblLoss$ can be written in a form resembling the
payoff formulation \eqnref{likelihood} for the feature likelihoods.

First we invoke the independence established in \eqnref{depth-indep}:
\begin{equation}
  \DepthLoss(\EstSeam,\Seam)
  =
  \frac{1}{N}
  \sum_{x=1}^{\Width}
  \sum_{y=1}^{\Height}
  \frac
      {\abs{\tilde{\Depth}(x,y;\hat{\seam}_x) - \Depth(x,y;\Seam)}}
      {\Depth(x,y;\Seam)} ~.
\end{equation}
Defining a real matrix $\LossMat_{\Seam}$,
\begin{equation}
  \LossMat_{\Seam}(x,j) =
  \sum_{y=1}^H
  \frac
      {\abs{\tilde{\Depth}(x,y;j) - \Depth(x,y;\Seam)}}
      {\Depth(x,y;\Seam)} ~,
\end{equation}
we see that we can write $\DepthLoss$ in the form
\begin{equation}
  \DepthLoss(\EstSeam,\Seam)
  =
  \frac{1}{N}
  \sum_{x=1}^{\Width} \LossMat_{\Seam}(x,\hat{\seam}_x) ~.
  \label{eq:depth-loss-matrix}
\end{equation}
There is a similar form for $\LblLoss$, which we omit here due to
space constraints.

\subsubsection{Choosing a Loss Function}

Neither of the above losses is unequivocally the ``correct'' loss; the
choice will depend on the application. One might expect a strong
correlation between the losses, and indeed one can show analytically
that
\begin{equation}
  \DepthLoss(\EstSeam,\Seam) = 0
  \iff
  \LblLoss(\EstSeam,\Seam) = 0 ~.
\end{equation}
However, in our experiments we found only a weak correlation between
these losses away from the origin. For example, the scatter plot shown
in figure \figref{comparative-scatter} shows a significant number of
outliers that score very well on $\DepthLoss$ but poorly on
$\LblLoss$, and vice versa.

@@

We turn now to the problem of learning within the model described
above. Our learning task is to identify a prediction function
$\Predictor$ mapping observed features $\Features$ to reconstructions
$\Seam$. We seek the loss minimizer
\begin{equation}
  \OptPredictor = 
  \argmin_{\Predictor} 
  \Expected \Bigl[ 
    \Loss\bigl(\Predictor(\Features),\Seam\bigr)
    \Bigr] ~,
\end{equation}
which we approximate in the framework of empirical risk minimization as
\begin{equation}
  \OptPredictor = 
  \argmin_{\Predictor} 
  \sum_k \Loss\bigl(\Predictor(\Features_k),\Seam_k\bigr) ~,
\end{equation}
where $k$ indexes a training set. To perform this optimization we
turn to the tools of structured prediction \cite{Bakir07}, and in
particular the structured SVM \cite{Tsochantaridis04}. First we need
to define $\Predictor$. In this paper we consider predictors of the
form
\begin{equation}
  \Predictor_{\Model}(\Features) = \argmax_{\Seam} 
  \bigl\langle \Model ~,~ \JointFtr(\Features,\Seam) \bigr\rangle
  ~.
  \label{eq:predictor-class}
\end{equation}
Comparing with \eqnref{log-linear-posterior} we see that each
predictor of the form \eqnref{predictor-class} is simply implementing
MAP inference under some set of hyper--parameters $\Model$. We now
turn to the optimisation problem itself. Following the standard
approach \cite{Bakir07} we cast the learning problem as a constrained
optimisation problem,
\begin{equation}
  \begin{split}
    \min_{\Model,\xi} &
      \hspace{2mm} 
    \frac{1}{2} \|\Model\|^2 +
      C \sum_{k=1}^n \xi_k\\
    s.t. & \hspace{2mm} \forall k, \Seam \neq \Seam_k:~
      \bigl\langle\Model, \JointFtr(\Features_k,\Seam_k)\bigr\rangle -
      \bigl\langle\Model, \JointFtr(\Features_k,\Seam)\bigr\rangle
      \geq
      \Loss(\Seam,\Seam_k) - \xi_k ~.
  \end{split}
  \label{eq:svm-problem}
\end{equation}
Tsochantaridis \etal \cite{Tsochantaridis04} described an algorithm
for solving this minimisation that is now used extensively within
machine learning and computer vision. To apply this algorithm here we
must solve two inference problems:
\begin{enumerate}
  \item{\textit{Prediction.} This is the maximization described in
    \eqnref{predictor-class}.}
  \item{\textit{Separation.} The algorithm described in
    \cite{Tsochantaridis04} requires a user--supplied procedure to
    find the ``most--violated constraint'' at each iteration. That is,
    \begin{equation}
      \argmax_{\Seam}
        \bigl\langle\Model, \JointFtr(\Features_k,\Seam)\bigr\rangle
        + \Loss(\Seam,\Seam_k) ~.
        \label{eq:separation-problem}
    \end{equation}
  }
\end{enumerate}

Our solutions to both of the above build on the algorithm presented by Flint
\etal \cite{Flint10eccv,Flint11}, which is a dynamic programming
solution to problems of the form
\begin{equation}
  \argmax_{\Seam}
  \sum_x \PixelPayoff(x,\seam_x) - \sum_j \CornerPenalty(j;\Seam) ~.
  \label{eq:inference-problem}
\end{equation}

\subsection{Inference (Prediction)}

We showed in \secref{model} that \eqnref{inference-problem} can be
written in the form \eqnref{predictor-class}, so the prediction
problem is a straightforward application of \cite{Flint10eccv}. This
is as expected since, as we have already remarked,
\eqnref{predictor-class} is equivalent to MAP inference on indoor
Manhattan reconstructions, which was precisely the subject
of \cite{Flint10eccv}.

\subsection{Loss--Augmented Inference (Separation)}

It turns out that the separation problem can also be solved using the
dynamic programming algorithm mentioned above, as the following
proposition shows.
